---
title: "Visualization 04 Challenge"
author: "Maya Gann-Bociek & Matthew Schuelke, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output:
  github_document: default
  html_notebook: default
---

## Introduction

This time you will try to create a bar plots to plot the average BP for each age decade, and add an error bar as the standard deviation. 

## Tasks
1. Load dependencies
2. Load in the dataset "nhanes_c3.csv" into your environment
3. Explore the data
4. Plot the data

## Load Dependencies

The following code loads the package dependencies for our analysis:

```{r package-load-1}

```

We now have the packages needed for our analyses.

## Load Data 

Load in data "nhanes_c3.csv"

```{r}

```

## Explore the Data

Print and summary the dataset to get familar with your data

```{r}

```

## Plot the Data

Create a jitter plot to plot BP against ageDecase, and map gender to color

```{r}

```

In this graph, try to plot the average BP of each ageDecade and 

* plot male and female seperately 
* add standard diviation error bar to each bar

```{r}

```
