---
title: "Visualization 04 Challenge Complete"
author: "Maya Gann-Bociek & Matthew Schuelke, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output:
  github_document: default
  html_notebook: default
---

## Introduction

This time you will try to create a bar plots to plot the average BP for each age decade, and add an error bar as the standard deviation. 

## Tasks
1. Load dependencies
2. Load in the dataset "nhanes_c3.csv" into your environment
3. Explore the data
4. Plot the data

## Load Dependencies

The following code loads the package dependencies for our analysis:

```{r package-load-1}
# tidyverse packages
library(dplyr)   
library(tidyr)   
library(ggplot2)  
library(readr)    
library(here)
library(skimr)
```

We now have the packages needed for our analyses.

## Load Data 

Load in data "nhanes_c3.csv"

```{r}
nhanes <- read_csv(here('data', 'nhanes_c3.csv'))
```

## Explore the Data

Print and summary the dataset to get familar with your data

```{r}
print(nhanes)
skim(nhanes)
```

## Plot the Data

Create a jitter plot to plot BP against ageDecase, and map gender to color

```{r}
ggplot(nhanes, aes(x = ageDecade, y = BP, color = gender)) + 
  geom_jitter() + 
  labs(x = "Age Decade", color = "Gender")
```

In this graph, try to plot the average BP of each ageDecade and 

* plot male and female seperately 
* add standard diviation error bar to each bar

```{r}
nhanes %>% group_by(ageDecade, gender) %>%
  summarise(BP_avg = mean(BP),
            std = sd(BP),
            n = n(),
            m = 1.96 * std / sqrt(n)) -> nhanes_summ

ggplot(nhanes_summ, aes(x = ageDecade, y = BP_avg, fill = gender)) + 
  geom_bar(stat="identity",   position = position_dodge(width = 0.90)) +
    geom_errorbar(aes(ymin = BP_avg - m, ymax = BP_avg + m), 
                width = 0.2,   position = position_dodge(width = 0.90)) + 
  labs(x = "Age Decade", y = "BP", fill = "Gender") +
  theme_classic()
```
